---
title: "Evaluating the potential for a growth and recruit overfishing when harvest is allowed under the minimum length limit"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true  
  word_document:  
    fig_width: 6.5  
    fig_height: 6.5 
    reference_docx: style.docx
bibliography: refs.bib
---  


<!--

rmarkdown::render("index.Rmd",output_format=c("html_document","word_document"))
to do
1. chi for vbgf in analysis.R
2. need a maturity function
3. params: in yml read in outputs
-->



```{r,echo=FALSE,warnings=FALSE, message=FALSE}
source("R/1_global.R")  	
source("R/2_functions.R")         
source("R/3_load.R")     
source("R/5_tables.R") 
source("R/6_figures.R")  	
#source("R/8_analysis.R")
```

$Y=i=12t=1i[C(t)iW(t)i]dt$ 	(1)

where c(t)iwas the age-specific cach, W(t)i was the weight at age, i was 
the maximum age, and $i$ indexed species. Equation 1 was numerically 
integrated given inputs described below to calculate yield for weight at 
age and catch at age equations {Colvin, 2013 #4635}. 


Catch at age
We modeled catch as the age-specific product of fishing mortality and 
natural mortality as 


$C(t)i=F(t)Rie-(Mtr,i)e-(M+F(t))(tr,i-tb,i)  e-(M+F(t))(t-tr,i)$,	(2)

where F(t)i was the age specific fishing mortality, $R_i$ was the species specific number of fish recruited to the cohort, Miwas the instantaneous natural mortality rate,  was the age that a fish recruits to the fishery, $t_{b,i}$ was the age a fish recruited to the entire fishery, $t$ was age, and $i$ indexed species. Fishing mortality varied with age as 

	(3)


Weight at age


## SPR

## Overfishing analysis

We used the YPR model to evaluate 3 proposed MLL, 254 mm, 279.4 mm, and 
304.80; varying fishing mortalities below the MLL (0, 0.01, 0.05, 0.1, 
0.2) and exploitation rates above the MLL from 0 to 0.9 in increments of 
0.05. By evaluating various fishing mortality below the MLL we were able 
to evaluate the potential effect of fishing mortality below the MLL, 
while still evaluating that effect that different length limits would 
have on yield in the system. This would also allow us to see when, if at 
all, growth overfishing could occur in the system. Once the required 
information: length-weight relationship, the VBGF, and annual 
exploitation were calculated we were able to simulate 5 different Yield 
per Recruit models, each assessing a different conditional fishing 
mortality below the MLL and the three options for the proposed length 
limit; 254, 279.4, 304.8 mm. We visually assessed the fishery for the 
potential for growth overfishing, by looking for decline in yield with 
increased effort. 


# Results

We captured 478 crappie, 360 Black Crappie and 118 White Crappie. Of the 
crappie caught, 154 Black Crappie and 62 White Crappie were retained and 
used to fit age-length and length-weight models. Size structure varied 
between Black and White Crappie and total length varied from 105 to 359 
mm and 100 to 364 mm respectively (Table 1). 


## Predictive relationships

Length-weight
* strong evidence for difference between species (intercept and slope), 
competing models were 6 or more AIC points away 


1. no species difference: AIC= -226.1
1. intercept varies by species: AIC= -260.1
1. intercept and slope varies between species: AIC= -266.0


Age-length
Length-fecundity


## Yield per Recruit analysis

### Growth overfishing

For each scenario, the current MLL of 254 mm had the greatest yield at 
the 41% exploitation rate. Yield decrease in every scenario with an 
increased MLL. As the conditional fishing mortality below the MLL 
increases we see a shift to the right on the exploitation axis was the 
254 mm length limit yield exceeds the 279.4 and 3014.8 mm length limit 
yield. However, for each scenario this point occurs under the 41% 
exploitation line that this fishery experiences. 


### Recruit overfishing

Spawning potential ratios decreased with increasing exploitation and 
decreasing MLL. 

* SPR decreased with increasing fishing mortality below the MLL
* 12" offered most protection from recruit overfishing with no scenarios exceeding 0.3
* 11" offered lots of protection for a range of fishing mortality and 
0.3 was not exceeded until u was higher than xxx. 
* 10" offered some protection but it only for u values < xxxx.


# Discussion

Crappie are an important sportfish in the United States (Maceina and 
Stimpert 1998). They are often managed for as the same species because 
anglers rarely differentiate between the two species (Mitzner 1991). For 
this study we evaluated the two species together because under the 
current regulations there is no difference in length limits or bag 
limits for Black and White Crappie. 

However, one species is often more abundant than the other in one system 
(Sammons et al. 2002). Assuming that Catch per Unit Effort in Lake 
Washington is linearly related to abundance, it indicates that there is 
more Black Crappie than White Crappie in the system. However, there is 
data that suggests that there may be a bias for Black Crappie when using 
trap nets, resulting in more Black Crappie being sampled that white 
crappie when relying solely on trap style nets for catch data (Sammons 
et al. 2002). It is also is shown that electrofishing may be more 
effective in capturing larger fish needed to obtain age and growth data, 
and that trap nets may be biased to smaller, younger fish (Sammons et 
al. 2002). 

Electrofishing in the spring could be done to add data, and improve the 
yield per recruit models. In our study, the mean length and weight was 
greater for Black Crappie than White Crappie. 

The mean age of White Crappie captured was also lower than the mean age 
of Black Crappie captured. This could suggest that trap nets are biased 
to capturing more Black Crappie, and that some electrofishing may also 
be needed to sample the larger fish in the system. 

Length limits have become increasingly popular ways on managing to meet 
angler demands of larger fish in the system. In many systems, harvest 
numbers, not larger sized fish, is of more importance to crappie anglers 
(Hale et all 1999). Lake Washington may be unique in that some anglers 
are specially seeking the larger trophy sized fish, yet others are 
seeking large numbers of fish harvested (N. Aycock, unpublished data). 
In a study by Hale et al. (1999), a 254 mm length limit was imposed on 
the Delaware Reservoir to increase size of crappie caught without 
greatly reducing harvest numbers. The length limit was successful in 
increasing the mean size of fish caught, but it did dramatically 
decrease annual yield. However, improvements to fisheries are specific 
to the study area and are susceptible to fluctuating year class strength 
(Webb and Ott 1991). 

In this study we noted that in every scenario presented yield would most 
likely decrease with an increased MLL. However, by increasing the length 
limit, the mean size of fish harvested would most likely increase. 
Keeping the “5 under rule” in place would still allow for the 
harvest of fish below the MLL, allowing anglers to still keep a 254 mm 
if, if the MLL length limit was increased. 

Miranda and Allen (1995) suggest that reducing exploitation by the use 
of harvest restrictions does not always improve yield. They found that 
yield is only likely to increase due to harvest restrictions when growth 
is fast and natural mortality is low. However, their model does suggest 
that decreasing annual exploitation, or restricting harvest, can 
increase the average weight of fish caught, but only when growth is fast 
and natural mortality is low. According to the models outlined in their 
paper the population in Lake Washington is a ??? growing population with 
(high/low natural mortality). Another issue often associated with 
managing crappie populations is how variable recruitment into crappie 
populations is. This may affect the estimate of total annual mortality 
of the population. A catch curve analysis was used to estimate our 
annual mortality; this method assumes that there is constant recruitment 
into the population (Allen 1997). Crappie populations have been shown to 
experience variability in recruitment to age-1 (McDonough and Buchanan 
1991; Webb and Ott 1991: Mitzner 1995). Our models do not take into 
account variable annual mortality rates, and yield from year to year 
will most likely fluctuate with fluctuating recruitment. 

White Crappie are thought to grow faster than Black Crappie due to their 
piscivorous diet (Ellison 1984). In a 2002 study, Sammons et al. found 
data that supported this conclusion. They found that by age 4, Black 
Crappie were on average 30 mm smaller than White Crappie. Variability in 
size at the same age could affect the age at which one of the species is 
affected by the MLL. If White Crappie are reaching the MLL before Black 
Crappie, and are less abundant that Black Crappie in the system, catch 
rates could be potentially affected. 

* Blacks versus whites and justification for combining, CPUE suggests 
they are a small component of Crappie stock, But there could be some 
habitat segregation (any papers documenting different growth? Different 
mortality rates?) 
* Effect of increasing MLL ? increased opportunities to harvest 5 under 
because more fish from 8? 12” versus 8? 10”, Creel indicates 8” 
over is harvested (N. Aycock Unpublished data) 


# Acknowledgments

We thank the Mississippi Department of Wildlife, Fisheries, and Parks 
for funding; Chad Washington, and Donte Green for leading fish 
collections and providing data; and Maddy Ruble for help aging otoliths. 



# Tables

Tables
Table 1. Model inputs
Parameter
Species
Value
Units
Source
Length-age ()
ln(a)
Black Crappie
-13.74
unitless
Estimated


White Crappie
-14.94
unitless
Estimated
b
Black Crappie
3.51
unitless
Estimated


White Crappie
3.71
unitless
Estimated
Growth ()
t0
Both
-0.85
yr
Estimated
Linf
Both
354.54
mm
Estimated
k
Black Crappie
0.35
yr^-1
Estimated


White Crappie
0.32
yr^-1
Estimated
Sexual maturity
Length at maturity
Both
180
mm
Froese and Binohlan 2000
Mortality and harvest
M
Black Crappie
66%
yr^-1
MDWFP


White Crappie
66%
yr^-1
MDWFP
F
Both
0.33
yr^-1
MDWFP
Minimum harvest size
Both
203.3
mm
MDWFP
Population characteristics
Species ratio
Black Crappie
0.25
unitless
Estimated


White Crappie
0.75
unitless
Estimated
Maximum age
Black Crappie
act=7 est= 7.72


Estimated


White Crappie
act=4 est=8.52


Estimated
Sex ratio
Both
0.5
unitless
Assumed


Figures

```{r,echo=FALSE, message=FALSE, warnings=FALSE}
figures(1)
```

Figure 1. Age (x-axis) and length (y-axis; mm) (top panel, Length 
(x-axis; mm) and Weight (y-axis; g) (middle panel), Length (x-axis; mm) 
and length (y-axis; mm) (bottom panel) of black and white crappie 
captured in Lake Washington, Mississippi. 

```{r,echo=FALSE, message=FALSE, warnings=FALSE}
figures(2)
```

Figure 2. Yield per 1000 recruits (y-axis) for varying conditional 
natural mortality rates and exploitation rates. Top panels represent no 
fishing mortality below the minimum length limit. The middle and bottom 
rows represent 0.1 and 0.2 respectively. 



```{r,echo=FALSE, message=FALSE, warnings=FALSE}
figures(3)
```

Figure 3. Spawning potential ratio (y-axis) for varying conditional 
natural mortality rates and exploitation rates. Top panels represent no 
fishing mortality below the minimum length limit. The middle and bottom 
rows represent 0.1 and 0.2 respectively. 



# References



